---
title: "Import AQI data"
author: "Michael Andreae"
date: "3 September 2015"
output: html_document
---

# Insurance status predicts antiemetic use 
We investigate the Hypothesis that socioeconomic status (**SES**) predicts anesthesia quality.

We use either insurance status or median income in patient address zip code as markers of socioeconomic status and  antiemetic use as marker of anesthesia quailty. We use the population in the Public Use File *NACOR*, the National Anesthesia Clinical Outcome Registry, of the Anesthesia Quality Institute *AQI* with electronic anesthesia records recording antiemetic use.


```{r, packages, message=FALSE, echo=FALSE, warning=FALSE}
require(knitr) # required to set options in Markdown R
```

```{r, global_options, echo=FALSE}
# set options
opts_chunk$set(fig.width=6, fig.height=4, fig.path='Figs/',
               echo=FALSE, warning=FALSE, message=FALSE)
# set working directory to source file location
#setwd("C:/Users/Micheal/Dropbox/Professional/AQI/AQI_Import")
```

# Import Data and generate clean data files

```{r, eval=FALSE, echo=TRUE}
# run only once

setwd("C:/Users/Micheal/Dropbox/Professional/AQI/Robert/Data/")
PUF <- read.csv("Data/PUF.csv")

setwd("C:/Users/Micheal/Dropbox/Professional/AQI/AQI_Import/Data")
save(PUF, file="PUF.Rdata")
rm(list = ls())
```

## Load Rdata AQI raw dataset 

```{r}
rm(list = ls())
# set working directory to source file location
#setwd("C:/Users/Micheal/Dropbox/Professional/AQI/AQI_Import")
load("Data/PUF.Rdata")
```

```{r countingNA}
n <- rep(0,20)
myNAs <- rep(0,20)
nCases<- 1 
nPayment<- 2
nSurgical_cpt <- 3
nDuration<- 4
nDuration <- 5
nAge <-6
nAge_group <- 7
nSex <- 8
nInpatient <- 9
nIncome_Median <- 10
nAnesType <- 11
nProcedure <- 12
nCase_type <-13
nASA_imputed <- 14
nASA <-15
nProv <- 16
n[nCases] <- length(PUF$caseID)
```

## Median income versus insurance status as predictors of socioeconomic status

### Median income

```{r}
summary(PUF$Income_MedianR)
hist(PUF$Income_MedianR)
myNAs[nIncome_Median] <-sum(is.na(PUF$Income_MedianR))
```

A predictor of socioeconomic status could be the median income in the zip code of the patients home address, coded *Income_MedicanR* , a `r class(PUF$Income_MedianR)` with `r myNAs[nIncome_Median]` NA values. We will generate a separate file with median income as predictor at the bottom. 

### Insurance status

```{r}
PUF$Payment[PUF$Payment == ""] <- NA
myNAs[nPayment] <-sum(is.na(PUF$Payment))

PUF$Payment <- droplevels(PUF$Payment)
levels(PUF$Payment)[1] <- "Commercial"
```

The predictor insurance status (*Payment*) is coded in `r length(levels(PUF$Payment))` levels as `r levels(PUF$Payment)`, and has `r myNAs[nPayment] ` cases with without insurance information, *NA*

### Missing data on insurance status and median income

```{r}
OverlapPayment_Income_MedianR <- table(is.na(PUF$Payment),
                                       is.na(PUF$Income_MedianR))
colnames(OverlapPayment_Income_MedianR) <- c("Observed Income", "NA")
rownames(OverlapPayment_Income_MedianR) <- c("Observed Insurance", "NA")

kable(OverlapPayment_Income_MedianR)
```

There is considerable overlap in the missingness of information on *Income_MedicanR* and insurance status coded as *Payment*. In other words, data are missing in different subpopulatons for *Income_MedianR* and *Payment*. This makes it feasible to test our hypothesis in two different subsets of the NACOR database

* The population with complete data on median income at patient address, using *Income_MedianR* as predictor for poor healthcare quality. 

* The population with complete data on insurance status, using insurance status coded as *Payment* as predictor for poor healthcare quality. 

**Summary:**
We generate several clean data files.
The first set of files uses insurance status as predictor;
the subsequent set of files may use median income (in patient address zip code) as predictor.

# Predictor: insurance status

We generate the first set of files; We limit the analysis on the complete cases with *Payment* information on insurance status.

```{r Payment}
PUF$Payment[PUF$Payment == ""] <- NA
PUF$Payment <- droplevels(PUF$Payment)
levels(PUF$Payment)[1] <- "Commercial"
myNAs[nPayment] <- sum(is.na(PUF$Payment))
```

We start out with `r n[nCases]` individual anesthesia records. The predictor insurance status (*Payment*) is coded in `r length(levels(PUF$Payment))` levels as `r levels(PUF$Payment)`, we found `r myNAs[nPayment]` cases without insurance information, (orginally coded as "").

## Outcome variables: antiemetic administration

We focus on the antiemetics *ondansetron*, *dexamethason* and *droperidol*, the only agents with convincing evidence for effect. We code antiemetic administrations logical in *ondan* and as factor in *ondansetron*

```{r Ondansetron}
PUF$ondan <- 
  (PUF$Antiemetics_ondansetron)==1
PUF$ondansetron <- 
  as.factor(PUF$Antiemetics_ondansetron)
levels(PUF$ondansetron) <- c("no Ondan", "Ondan")
```

```{r Dexamethason}
PUF$dex <- 
  (PUF$Antiemetics_dexamethason)== 1
PUF$dexamethason <- 
  as.factor(PUF$Antiemetics_dexamethason)
levels(PUF$dexamethason) <- c("no Dex", "Dex")
```

```{r Droperidol}
PUF$drope <- 
  (PUF$Antiemetics_droperidol)==1
PUF$droperidol <- 
  as.factor(PUF$Antiemetics_droperidol)
levels(PUF$droperidol) <- c("no Drope", "Drope")
```

### Combined antiemetic administration

```{r}
kable(table(PUF$ondansetron, 
            PUF$dexamethason),
            caption = "Cases with Ondansetron versus Dexamethason")

kable(table(PUF$ondansetron, 
            PUF$droperidol),
            caption = "Cases with Ondansetron versus Droperidol")

kable(table(PUF$dexamethason, 
            PUF$droperidol),
            caption = "Cases with Dexamethason versus Droperidol")
```

```{r}
PUF$either <- PUF$ondan | PUF$dex
PUF$ondan_dex_either <- 
  as.factor((PUF$ondan == 1) | (PUF$dex ==1))
levels(PUF$ondan_dex_either) <- c("neither", "either")
```

Both antiemetics ondansetron **OR** dexamethason were sometimes administered together:

This is coded in *ondan_dex_either* and in *either*

## Potential confounders and other variables

### procedures

```{r surgical_cpt}
levels(PUF$surgical_cpt)[1] <- "NA"
PUF$surgical_cpt <- droplevels.factor(PUF$surgical_cpt)
myNAs[nSurgical_cpt] <- sum(is.na(PUF$surgical_cpt))
kable(table(is.na(PUF$surgical_cpt), is.na(PUF$Payment)))
PUF$cpt <- PUF$surgical_cpt 
```

There are `r nlevels(PUF$surgical_cpt)` unique *surgical_cpt* codes, but unfortunately `r myNAs[nSurgical_cpt]` are NA, which compounds the missingness when tabulated with NAs in *Payment* above.

### practice ID versus facility ID

```{r}
PUF$practice <- as.factor(PUF$practiceID)
levels(PUF$practice) <- c("A", "B", "C", "D", "E", "F")

PUF$facility <- as.factor(PUF$facilityID)
levels(PUF$facility) <- c("alpha", "beta", "gamma", 
                          "epsilon", "kappa", "sigma",
                          "rho", "tau", "omega")
```

We recode the practices and facilities to A through F to prevent reidentification.

```{r}
kable(table(PUF$facility,PUF$practice))
```

The table of facility ID versus practice ID suggests that five practices have only one facility ID and one practice (=5013437) has three (sub) facilities. We will simplify by using practice ID, which has no NA.




### case_duration_minutes

```{r, echo=TRUE}
PUF$case_duration_minutes[PUF$case_duration_minutes==-1] <- NA
myNAs[nDuration] <- sum(is.na(PUF$case_duration_minutes))
hist(PUF$case_duration_minutes)
```

Case duration in minutes (*case_duration_minutes*) is an integer and has `r myNAs[nDuration] ` missing values coded as *NA*.

### patient age



```{r, echo=TRUE}
PUF$patient_age[PUF$patient_age==-1] <- NA
myNAs[nAge] <-sum(is.na(PUF$patient_age))

hist(PUF$patient_age, 
     main = "Histogram of Age Distribution",
     xlab = "Age in years")
```

Patient age (*patient_age*) is an integer with a distribution below and has `r myNAs[nAge]` missing values coded as NA.

### patient_age_group

```{r, echo=TRUE}
levels(PUF$patient_age_group)[2] <- "1-18"
levels(PUF$patient_age_group)[1] <- NA
myNAs[nAge_group] <-sum(is.na(PUF$patient_age_group))

PUF$patient_age_group <- 
  relevel(PUF$patient_age_group, "Under 1")
PUF$patient_age_group <- relevel(PUF$patient_age_group, ref = "19 - 49")
```

Patient age group (*patient_age_group*) is a factor with `r length(levels(PUF$patient_age_group))` levels: `r levels(PUF$patient_age_group)`; it has `r myNAs[nAge_group]` missing values initially coded as -1, which we recoded as NA and removed from the dataset. (We corrected the miscoding from "18-Jan"" to "1-18").

### patient_sex

```{r}
PUF$patient_sex <- 
  as.factor(PUF$patient_sex)

levels(PUF$patient_sex)[1] <- NA
levels(PUF$patient_sex)[1] <- "female"
levels(PUF$patient_sex)[2] <- "male"

myNAs[nSex] <- sum(is.na(PUF$patient_sex))
```

Patient gender (*patient_sex*) is recoded as factor with the two levels `r levels(PUF$patient_sex)` and `r myNAs[nSex]` NAs, which are removed from the dataset.

### in_or_out_patient

```{r}
PUF$in_or_out_patient <- 
  as.factor(PUF$in_or_out_patient)

levels(PUF$in_or_out_patient)[1] <- NA
levels(PUF$in_or_out_patient)[1] <- "Outpatient"
levels(PUF$in_or_out_patient)[2] <- "Inpatient"

myNAs[nInpatient] <- sum(is.na(PUF$in_or_out_patient))
```


in- or outpatient status (*in_or_out_patient*) is recoded as a factor with the two levels `r levels(PUF$in_or_out_patient)` and `r myNAs[nInpatient]` NAs, which are too numerous to include this variable in the dataset.

### primary_anesthesia_type

```{r}
PUF$primary_anesthesia_type <- 
  as.factor(PUF$primary_anesthesia_type)
levels(PUF$primary_anesthesia_type)[1] <- NA
levels(PUF$primary_anesthesia_type) <- c("General", 
                                                "Neuroaxial", 
                                                "Regional", 
                                                "MAC", 
                                                "Sedation", "Local" , 
                                                "Other")
kable(mytable <- data.frame(n=summary(PUF$primary_anesthesia_type)))
myNAs[nAnesType] <- sum(is.na(PUF$primary_anesthesia_type))
```

*primary_anesthesia-type* is recoded as a factor with  `r length(levels(PUF$primary_anesthesia_type))` levels [`r levels(PUF$primary_anesthesia_type)`] and `r myNAs[nAnesType]` cases at this time do not have a *primary_anesthesia_type* code defined. We consider to control with a fixed or a random effect for *primary_anesthesia_type* code. 

```{r}
levels(PUF$primary_anesthesia_type)[5:7] <- NA
```

We did exclude other, sedation and local leaving us with `r myNAs[nAnesType]` NA anesthesia_type cases cases.

### procedure_status

```{r}
PUF$procedure_status <- 
  as.factor(PUF$procedure_status)
levels(PUF$procedure_status)[1] <- NA
levels(PUF$procedure_status) <- c("Emergency", "Elective")
myNAs[nProcedure] <- sum(is.na(PUF$procedure_status))
```

It would make sense to try to control for *procedure_status*, (which indicates if the case was `r levels(PUF$procedure_status) [1]` or `r levels(PUF$procedure_status)[2]`); but `r myNAs[nProcedure]` of the cases do not have a *procedure_status* code defined, which obviously are too many to exclude.

### case_type

```{r}
# load("Analysis/Data/PUF.Rdata")

PUF$case_type <- 
  as.factor(PUF$case_type)
levels(PUF$case_type)[1] <- NA
myNAs[nCase_type] <- sum(is.na(PUF$case_type))
```

It would make sense to try to control for *case_type*, (which indicates if the case was `r levels(PUF$case_type)` ..., but `r myNAs[nCase_type]` of the cases do not have a *case_type* code defined, which obviously are too many to exclude.


### asaps_imputed

```{r}
PUF$asaps_imputed <- 
  as.factor(PUF$asaps_imputed)
levels(PUF$asaps_imputed)[1] <- NA
myNAs[nASA_imputed] <- sum(is.na(PUF$asaps_imputed))
```

### asaps

```{r}
PUF$asaps <- 
  as.factor(PUF$asaps)
levels(PUF$asaps)[7] <- NA # exclude the ASA 6
levels(PUF$asaps)[1] <- NA
myNAs[nASA] <- sum(is.na(PUF$asaps))
```

It would make sense to try to control for *asaps* or *asaps_imputed*, (ASA Status, which indicates how sick a patient was, we should also exclude ASA status 6 (organ donars) and only `r myNAs[nASA]` of the remaining cases do not have an ASA status recorded; so we exclude them, leaving us with `r dim(PUF)[1]` unique cases with also *asaps* as a predictor.


### prov1

```{r}
PUF$prov1 <- 
  as.factor(PUF$prov1)
levels(PUF$prov1)[1] <- NA
myNAs[nProv] <- sum(is.na(PUF$prov1))
```


It would be great to control for individual provider behavior, to show variability among providers in their propensity to admister antiemetics contingent on insurance status. There are `r length(levels(PUF$prov1))` different *prov1* levels, I believe they are coding for individual providers. `r myNAs[nProv]` of the remaining cases do not have the *prov1*  recorded; 

### Median Income of patient zip code

```{r}
IncomeSummary <- summary(PUF$Income_MedianR)
myNAs[nIncome_Median] <- sum(is.na(PUF$Income_MedianR))

library(dplyr)
library(knitr)


hist(PUF$Income_MedianR, main = "Histogram of Median Income", xlab = "Median income in patients home zip code in $", breaks =40)

```

The above histogram show the distribution of median incomes in the patients home zip codes, used as a indicator of SES. *Income_MedianR*, an integer, encodes for the median income in the patient home address zipcode. Median Income is `r summary(PUF$Income_MedianR)[2]`, with an IQR [`r summary(PUF$Income_MedianR)[c(2,5)]`]. There are  `r myNAs[nIncome_Median]` NAs. We will scale the variable *Income_MedianR* to units of income of $100,000  to make the regression coefficient more interpretable. 1 correspondes therefore to a median income of $100,000 in the patients home zip code area. The median of the zip code income medians is `r median(PUF$Income_MedianR, na.rm= TRUE)`

#### Deviding into median income quantile
```{r}
PUF$IncomeQ <- cut (PUF$Income_MedianR, 
                     quantile(PUF$Income_MedianR, na.rm = TRUE),
                     labels =c("very low","low", "middle", "high")
  )
```

We break the data into four income quantiles with the cut points:
`r quantile(PUF$Income_MedianR, na.rm = TRUE)`. 

## Table of missing patterns

```{r}
# which position in the PUF datafram do the variable have
iAgeGroup <- 15
iPatientSex <-16
iASA <- 17
iAnes_type <- 28
isurgicalCPT <- 29
iProvider <- 62
iIncome_MedianR <- 76
iPayment <- 78

# make a list to call them in the loop
ilist<- c(iProvider,
          isurgicalCPT,
          iPayment,
          iIncome_MedianR,
          iASA,
          iAnes_type,
          iPatientSex,
          iAgeGroup)

# how many variables are there
n_i <- length (ilist)

# define dimensions for a table of missingness
mtable <- matrix(rep(0,n_i^2), ncol=n_i)


# for each variable fill the diagonal with the number of missing NAs
for (d in 1:n_i) {  mtable[d,d] <- sum( is.na(PUF[,ilist[d] ] )) }

for (i in 1:(n_i-2)) {
  
  # for each other variable
  for (j in (i+1):(n_i)) {
    # make a the joint missingness table
    overlap <- table(is.na(PUF[,ilist[i] ]),
                     is.na(PUF[,ilist[j] ]))

    # put the number of units with both variables observed in upper triangle                   
    mtable[i,j] <- overlap[1,1]
    
    # put the number of units with neither variables observed in lower triangle         
    mtable[j,i] <- overlap[2,2]
  }
}

# Do for last comparison i+1 and j= n_i
    overlap <- table(is.na(PUF[,ilist[7] ]),
                     is.na(PUF[,ilist[8] ]))

    # put the number of units with both variables observed in upper triangle                   
    mtable[(i+1),n_i] <- overlap[1,1]
    
    # put the number of units with neither variables observed in lower triangle         
    mtable[n_i,(i+1)] <- overlap[2,2]

# label table columns and rows

colnames(mtable) <- rownames(mtable) <- c("Provider",
          "CPT",
          "Insurance",
          "Income Median",
          "ASA",
          "Anes_type",
          "Gender",
          "Age Group")
missingnesstable <- mtable
save(missingnesstable, file = "Results/missingnesstable.Rdata")
kable(mtable, caption = "Table of missing patterns. We tabulate in the diagonale the number of missing observations for ther variable; in the upper triagle of the table, the number of jointly observed and in the lower triangle the number of jointly missing observations for the corresponding variables." )
```


# Clean AQI_cpt dataframe saved as Rdata

```{r, eval=TRUE}

attach(PUF)

myAQI_cpt<- data.frame(ond= ondansetron,
                       ondL=ondan,
                      
                       dexa= dexamethason,
                       dexL= dex,
                       
                       dro= droperidol,
                       droL= drope,
                       
                       any = ondan_dex_either,
                       either = either,
                       
                       pay = Payment,
                       income = Income_MedianR/ 100000, # scale to $100,000 income units
                       incomeQ = IncomeQ, # income quantiles
                       age = patient_age, 
                       age_group = patient_age_group, 
                       sex = patient_sex,
                       
                       cpt = cpt,
                       ASA = asaps,
                       anes_type = primary_anesthesia_type,
                       
                       practice = practice,
                       prov = prov1
                       )
save (myAQI_cpt, file="myAQI_cpt.Rdata")
str(myAQI_cpt)
```

The dataset has dimension `r (dim.data.frame((myAQI_cpt)))`, before we limit to only complete cases.

## Attrition flow diagram

```{r}

load("myAQI_cpt.Rdata")
#counters of remaining observations and cases lost after removing variable
nObservations <- rep(0, 10)
nlost <- rep(0, 10)

# Total number of cases in PUF
nObservations[1] <- 8000000
nlost[1] <- 0

# Number of cases in customized dataset
nObservations[2] <- 441645 # anesthesia cases in PUF
nlost[2] <- 8000000 - 441645


# Number of Complete cases after sequentially removing missing data for various variables 

# prov
nlost[3] <- sum(is.na(myAQI_cpt$prov))
reducedAQI <- myAQI_cpt[complete.cases(myAQI_cpt$prov),]
nObservations[3] <- dim(reducedAQI)[1]

# cpt
nlost[4] <- sum(is.na(reducedAQI$cpt))
reducedAQI <- reducedAQI[complete.cases(reducedAQI$cpt),]
nObservations[4] <- dim(reducedAQI)[1]

# pay
nlost[5] <- sum(is.na(reducedAQI$pay))
reducedAQI <- reducedAQI[complete.cases(reducedAQI$pay),]
nObservations[5] <- dim(reducedAQI)[1]

# income
nlost[6] <- sum(is.na(reducedAQI$income))
reducedAQI <- reducedAQI[complete.cases(reducedAQI$income),]
nObservations[6] <- dim(reducedAQI)[1]

# ASA
nlost[7] <- sum(is.na(reducedAQI$ASA))
reducedAQI <- reducedAQI[complete.cases(reducedAQI$ASA),]
nObservations[7] <- dim(reducedAQI)[1]

# Anest_type
nlost[8] <- sum(is.na(reducedAQI$anes_type))
reducedAQI <- reducedAQI[complete.cases(reducedAQI$anes_type),]
nObservations[8] <- dim(reducedAQI)[1]

# Gender
nlost[9] <- sum(is.na(reducedAQI$sex))
reducedAQI <- reducedAQI[complete.cases(reducedAQI$sex),]
nObservations[9] <- dim(reducedAQI)[1]

# Age group
nlost[10] <- sum(is.na(reducedAQI$age_group))
reducedAQI <- reducedAQI[complete.cases(reducedAQI$age_group),]
nObservations[10] <- dim(reducedAQI)[1]

varnames <-  c("PUF",
          "Customized Dataset",
          "Complete Provider Data",
          "Complete surgical CPT codes",
          "Complete Insurance Information",
          "Complete Income Median Information",
          "Complete ASA Classification",
          "Complete Anesthesia type data",
          "Complete Gender Information",
          "Complete Age Information")
attritionTable <- data.frame(Dataset= varnames, lost=nlost, remaining = nObservations)
save(attritionTable, file="results/AttritionTable.Rdata")
kable(attritionTable, caption = "Flow diagram shoing remaining number of observations after sequentially removing cases with incomplete data on certain variables. The main loss occurs by limiting the Public User File to cases with antiemetic medication data.")
```

## Complete case dataset ccAQI_cpt

```{r limitCompleteCases}
ccAQI_cpt <- myAQI_cpt[complete.cases(myAQI_cpt),]
save (ccAQI_cpt, myAQI_cpt, file="myAQI_cpt.Rdata")
str(ccAQI_cpt)
