---
title: "Import AQI data"
author: "Michael Andreae"
output: html_document
---

#Insurance status predicts antiemetic use 
We investigate the Hypothesis that insurance status predicets antiemetic use the population in the Public Use File of the Anestehsia Quality Institute with electronic anesthesia records recording antiemetic use controlling for procedure.



```{r, packages, message=FALSE, echo=FALSE, warning=FALSE}
require(knitr) # required to set options in Markdown R
```

```{r, global_options, echo=FALSE}
# set options
opts_chunk$set(fig.width=6, fig.height=4, fig.path='Figs/',
               echo=FALSE, warning=FALSE, message=FALSE)
setwd("C:/Users/Micheal/Dropbox/Professional/AQI/AQI_Import")
```

# Load original dataset and save as Rdata

```{r, eval=FALSE, echo=TRUE}
setwd("C:/Users/Micheal/Dropbox/Professional/AQI/Robert/Data/")
PUF <- read.csv("PUF.csv")

setwd("C:/Users/Micheal/Dropbox/Professional/AQI/AQI_Import/Data")
save(PUF, file="PUF.Rdata")
rm(list = ls())
```

# Load Rdata AQI raw dataset

```{r}
rm(list = ls())
setwd("C:/Users/Micheal/Dropbox/Professional/AQI/AQI_Import")
load("Data/PUF.Rdata")
```


```{r countingNA}
attach(PUF)
n <- rep(0,20)
myNAs <- rep(0,20)
nCases<- 1 
nPayment<- 2
nSurgical_cpt <- 3
nDuration<- 4
nDuration <- 5
nAge <-6
nAge_group <- 7
nSex <- 8
nInpatient <- 9
n[nCases] <- length(caseID)
```

## Predictor insurance status

```{r Payment}
attach(PUF)
length(PUF$Payment[Payment == ""])
PUF$Payment[PUF$Payment == ""] <- NA
PUF$Payment <- droplevels(PUF$Payment)
levels(PUF$Payment)[1] <- "Commercial"
myNAs[nPayment] <- sum(is.na(PUF$Payment))
```

We start out with `r n[nCases]` individual anesthesia records. The predictor insurance status (*Payment*) is coded in `r length(levels(PUF$Payment))` levels as `r levels(PUF$Payment)`, we found `r myNAs[nPayment]` cases without insurance information, (orginally coded as "").

## Outcome variable antiemetic administration

We code antiemetic administration logical in *ondan* and as factor in *ondansetron*

### Ondansetron

```{r}
attach(PUF)
PUF$ondan <- 
  (Antiemetics_ondansetron)==1
PUF$ondansetron <- 
  as.factor(Antiemetics_ondansetron)
levels(PUF$ondansetron) <- c("not admin", "admin")
```

### Dexamethason

```{r}
attach(PUF)
PUF$dex <- 
  (Antiemetics_dexamethason)==1
PUF$dexamethason <- 
  as.factor(Antiemetics_dexamethason)
levels(PUF$dexamethason) <- c("not admin", "admin")
```

### Droperidol

```{r}
attach(PUF)
PUF$drope <- 
  (Antiemetics_droperidol)==1
PUF$droperidol <- 
  as.factor(Antiemetics_droperidol)
levels(PUF$droperidol) <- c("not admin", "admin")
```

### Ondansetron or Dexamethason

```{r}
attach(PUF)
PUF$either <- ondan | dex
PUF$ondan_dex_either <- 
  as.factor((PUF$ondan == 1) | (PUF$dex ==1))
levels(PUF$ondan_dex_either) <- c("neither", "either")
```

```{r}
kable(table(PUF$ondansetron, PUF$dexamethason))
table(PUF$ondan_dex_either)
```
Both antiemetics ondansetron **OR** dexamethason were sometimes administered together:

This is coded in *ondan_dex_either* and in *either*

## Potential confounders and other variables

### procedures

```{r surgical_cpt}
levels(PUF$surgical_cpt)[1] <- "NA"
PUF$surgical_cpt <- droplevels.factor(PUF$surgical_cpt)
myNAs[nSurgical_cpt] <- sum(is.na(PUF$surgical_cpt))
kable(table(is.na(PUF$surgical_cpt), is.na(PUF$Payment)))
PUF$cpt <- PUF$surgical_cpt 
```

There are `r nlevels(surgical_cpt)` unique *surgical_cpt* codes, but unfortunately `r myNAs[nSurgical_cpt]` are NA, which compounds the missingness when tabulated with NAs in *Payment* above.

### practice ID versus facility ID

```{r}
attach(PUF)
kable(table(facilityID,practiceID))
```

The table of facility ID versus practice ID suggests that five practices have only one facility ID and one practice (=5013437) has three (sub) facilities. We will simplify by using practice ID, which has no NA.


### case_duration_minutes

Case duration in minutes (*case_duration_minutes*) is an integer and has `r sum(case_duration_minutes==-1)` missing values coded as -1, which we recoded as NA and removed from the dataset.

```{r, echo=TRUE}
PUF$case_duration_minutes[PUF$case_duration_minutes==-1] <- NA
myNAs[nDuration] <- sum(is.na(PUF$case_duration_minutes))
```


### patient age

Patient age (*patient_age*) is an integer with a distribution below and has `r sum(patient_age==-1)` missing values coded as -1, which we recoded as NA and removed from the dataset

```{r, echo=TRUE}
PUF$patient_age[PUF$patient_age==-1] <- NA
myNAs[nAge] <-sum(is.na(PUF$patient_age))

hist(PUF$patient_age, 
     main = "Histogram of Age Distribution",
     xlab = "Age in years")
```


### patient_age_group

```{r, echo=TRUE}
levels(PUF$patient_age_group)[2] <- "1-18"
levels(PUF$patient_age_group)[1] <- NA
myNAs[nAge_group] <-sum(is.na(PUF$patient_age_group))

PUF$patient_age_group <- 
  relevel(PUF$patient_age_group, "Under 1")
```

Patient age group (*patient_age_group*) is a factor with `r length(levels(PUF$patient_age_group))` levels: `r levels(PUF$patient_age_group)`; it has `r myNAs[nAge_group]` missing values initially coded as -1, which we recoded as NA and removed from the dataset. (We corrected the miscoding from "18-Jan"" to "1-18").

### patient_sex

```{r}
PUF$patient_sex <- 
  as.factor(PUF$patient_sex)

levels(PUF$patient_sex)[1] <- NA
levels(PUF$patient_sex)[1] <- "female"
levels(PUF$patient_sex)[2] <- "male"

myNAs[nSex] <- sum(is.na(PUF$patient_sex))
```

Patient gender (*patient_sex*) is recoded as factor with the two levels `r levels(PUF$patient_sex)` and `r myNAs[nSex]` NAs, which are removed from the dataset.

### in_or_out_patient

```{r}
PUF$in_or_out_patient <- 
  as.factor(PUF$in_or_out_patient)

levels(PUF$in_or_out_patient)[1] <- NA
levels(PUF$in_or_out_patient)[1] <- "Outpatient"
levels(PUF$in_or_out_patient)[2] <- "Inpatient"

myNAs[nInpatient] <- sum(is.na(PUF$in_or_out_patient))
```


in- or outpatient status (*in_or_out_patient*) is recoded as a factor with the two levels `r levels(PUF$in_or_out_patient)` and `r myNAs[nInpatient]` NAs, which are too numerous to include this variable in the dataset.

# Clean AQI_cpt dataframe saved as Rdata

```{r, eval=TRUE}

attach(PUF)
myAQI_cpt <- data.frame(Payment, 
                         ondan,
                         ondansetron,
                         dex,
                         dexamethason,
                         drope,
                         droperidol,
                         either,
                         ondan_dex_either,
                         cpt,
                         practiceID=as.factor(practiceID), 
                         case_duration_minutes, 
                         patient_age, patient_age_group, 
                         patient_sex)
myAQI_cpt <- myAQI_cpt[complete.cases(myAQI_cpt),]
(dim.data.frame((myAQI_cpt)))
save (myAQI_cpt, file="myAQI_cpt.Rdata")
str(myAQI_cpt)